\documentclass[12pt,a4paper]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english,arabic]{babel}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{titlesec}

\newcommand{\engnum}[1]{\textLR{#1}}

\setstretch{1.15}
\setlist[enumerate]{label=(\alph*), itemsep=0.4em, topsep=0.4em}
\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

\begin{document}

\begin{center}
    {\Large\bfseries الشروط والأحكام الخاصة بـ Slou Studio}
\end{center}

\vspace{1em}

\selectlanguage{arabic}

\section*{1. الحجز والتسجيل}
\begin{enumerate}
    \item تخضع جميع الحجوزات لتوفر الأماكن.
    \item يجب على العميلة إكمال عملية التسجيل وتقديم معلومات صحيحة ودقيقة ومحدّثة.
    \item يتم الحجز والشراء فقط من خلال القنوات والمنصات الرسمية المعتمدة من الاستوديو.
    \item يحتفظ الاستوديو بالحق في رفض أو إلغاء أي حجز أو عضوية أو باقة عند وجود مخالفة لهذه الشروط أو عند وجود سبب تشغيلي أو أمني أو صحي.
\end{enumerate}

\section*{2. الأسعار والدفع والضرائب}
\begin{enumerate}
    \item جميع الأسعار المعلنة تشمل ضريبة القيمة المضافة.
    \item قد تخضع الأسعار والرسوم والباقات والخدمات والمنتجات للتعديل أو التحديث في أي وقت دون إشعار مسبق.
    \item يجب سداد الرسوم كاملة وفق وسيلة الدفع المعتمدة قبل الاستفادة من الخدمة، ما لم يعلن الاستوديو خلاف ذلك.
    \item لا يجوز تقديم أي مبالغ أو هدايا مباشرة للمدربات أو الموظفات مقابل الخدمات، وتتحمل العميلة كامل المسؤولية عن أي مدفوعات شخصية خارج القنوات الرسمية.
\end{enumerate}

\section*{3. العضويات والباقات والحصص}
\begin{enumerate}
    \item جميع العضويات والباقات والحصص غير قابلة للاسترداد النقدي أو الاستبدال، وغير قابلة للتحويل إلى شخص آخر، ما لم يقرر الاستوديو خلاف ذلك كتابة.
    \item صلاحية الحصة الواحدة هي \engnum{7} أيام من تاريخ الشراء أو التفعيل، وصلاحية باقة \engnum{10} حصص هي \engnum{30} يوماً من تاريخ الشراء أو التفعيل.
    \item يجب استخدام الحصص والباقات قبل تاريخ الانتهاء المحدد، وأي حصص غير مستخدمة بعد تاريخ الانتهاء تُعد ساقطة ما لم يقرر الاستوديو خلاف ذلك.
    \item يحتفظ الاستوديو بحق تحديد عدد الجلسات أو الحصص أو الأنشطة لكل عضوية أو باقة بحسب الخطة المعلنة.
    \item يجوز حجز جلسة واحدة فقط في اليوم للشخص الواحد ما لم يعلن الاستوديو خلاف ذلك.
    \item الباقات أو الحصص أو العروض الخاصة تخضع للشروط والسعر الخاص المعلن لها.
\end{enumerate}

\section*{4. تحويل العضوية أو النقل أو التجميد}
\begin{enumerate}
    \item الأصل أن العضويات والباقات والحصص شخصية وغير قابلة للتحويل إلى شخص آخر.
    \item إذا أتاح الاستوديو خيار تحويل العضوية أو النقل بين الفروع أو الخطط، فيكون ذلك خاضعاً لموافقة الاستوديو والرسوم والشروط المعلنة، بما في ذلك أي عدد مجاني للتحويل وأي رسوم على التحويلات اللاحقة عند توفر حد أدنى من الحصص المتبقية.
    \item إذا أتاح الاستوديو تجميد العضوية أو الباقة، فيكون عدد مرات التجميد ومدته وفق نوع الباقة أو الخطة المعلنة وقت الشراء، وقد يشمل ذلك فترات تجميد محددة دون رسوم إضافية.
\end{enumerate}

\section*{5. الإلغاء وعدم الحضور والتأخر وقائمة الانتظار}
\begin{enumerate}
    \item يجب إلغاء الحجز قبل موعد الحصة بما لا يقل عن \engnum{6} ساعات.
    \item إذا تم الإلغاء ضمن المهلة المسموحة، فيجوز إعادة الحصة إلى الباقة أو الرصيد وفق السياسة المطبقة على نوع الحجز.
    \item الإلغاء المتأخر أو عدم الحضور يترتب عليه خصم الحصة أو احتساب كامل الرسوم بحسب نوع الخدمة والباقة، دون استرداد.
    \item في حال تكرار الإلغاء المتأخر أو عدم الحضور، يحق للاستوديو تعليق صلاحية الحجز مؤقتاً أو فرض قيود على الحجز، بما في ذلك تعليق الحجز لمدة محددة مثل \engnum{3} أيام أو أكثر وفق ما يراه الاستوديو مناسباً.
    \item يجب الحضور قبل بداية الحصة بوقت كافٍ.
    \item يسمح بدخول العميلة متأخرة بحد أقصى \engnum{15} دقيقة من وقت بدء الحصة، وبعد ذلك يحق للاستوديو رفض الدخول وتُحتسب الحالة كعدم حضور أو إلغاء متأخر دون استرداد.
    \item إذا كانت خدمة قائمة الانتظار متاحة، فإن الانضمام إليها قد يؤدي إلى إضافة العميلة تلقائياً للحصة عند توفر مكان، وتسري عليها عندئذٍ سياسة الإلغاء المعمول بها.
    \item تقع على العميلة مسؤولية إزالة اسمها من قائمة الانتظار أو متابعة حجوزاتها، وإذا اختارت الانضمام إلى أكثر من قائمة انتظار أو أكثر من جلسة، فإنها توافق على احتساب الحصص أو الرسوم وفق السياسة المطبقة عند تأكيد الحجز.
\end{enumerate}

\section*{6. الأهلية والعمر والهوية}
\begin{enumerate}
    \item العضوية والحصص الاعتيادية مخصصة لمن أعمارهن \engnum{16} سنة فأكثر، ما لم يعلن الاستوديو عن برنامج خاص بفئة عمرية مختلفة.
    \item إذا كان هناك برنامج مخصص للقاصرات أو الأطفال، فتطبق عليه شروطه الخاصة وموافقة ولي الأمر والحد الأدنى للعمر المعلن لذلك البرنامج.
    \item يجب على العميلة إبراز الهوية أو الإقامة أو أي وسيلة تحقق رقمية أو مستندات تعريف مطلوبة وفق الأنظمة أو وفق طلب الاستوديو.
    \item استخدام العضوية أو الحجز من قبل شخص غير مسجل أو غير مصرح له قد يؤدي إلى إلغاء العضوية أو الحجز فوراً دون تعويض.
\end{enumerate}

\section*{7. الصحة واللياقة والإفصاح الطبي}
\begin{enumerate}
    \item يجب على العميلة الحفاظ على تغطية تأمين صحي أو موافقة طبية عند الحاجة، وتتحمل مسؤولية التأكد من أهليتها الصحية للمشاركة.
    \item يجب على العميلة الإفصاح قبل المشاركة عن أي حالة صحية أو إصابة أو حمل أو عملية جراحية حديثة أو قيود بدنية أو أي أمر قد يؤثر في قدرتها على ممارسة التمارين بأمان.
    \item يجوز للاستوديو طلب موافقة أو تقرير طبي قبل السماح بالمشاركة في أي حصة أو نشاط.
    \item تؤكد العميلة أنها في حالة صحية مناسبة تسمح لها بالمشاركة في الأنشطة الرياضية، وأنها استشارت طبيبها عند الحاجة، خاصة عند وجود أمراض مزمنة أو ارتفاع ضغط الدم أو السكري أو الحمل أو أي ظروف صحية أخرى.
    \item يجب إبلاغ المدربات أو فريق الاستوديو فوراً عند الشعور بألم أو دوار أو ضيق تنفس أو أي أعراض غير معتادة قبل الحصة أو أثناءها أو بعدها.
    \item يحق للاستوديو إيقاف المشاركة أو رفضها مؤقتاً أو دائماً إذا رأى أن ذلك ضروري لصحة العميلة أو سلامتها أو سلامة غيرها.
    \item يحتفظ الاستوديو بحق تقديم الإسعافات الأولية الأساسية أو طلب المساعدة الطبية الطارئة عند الحاجة.
\end{enumerate}

\section*{8. المسؤولية وإخلاء المسؤولية}
\begin{enumerate}
    \item المشاركة في الحصص والأنشطة واستخدام المرافق والمعدات تكون على مسؤولية العميلة الشخصية.
    \item توافق العميلة على تحمل المخاطر المعتادة المرتبطة بالنشاط البدني والتمارين الرياضية، بما في ذلك الإجهاد أو السقوط أو الإصابات أو المضاعفات الصحية التي قد تنشأ أثناء التمرين أو بعده.
    \item لا يتحمل الاستوديو، بالقدر الذي يسمح به النظام، أي مسؤولية عن الإصابات أو الخسائر أو الأضرار الناتجة عن المشاركة أو عن مخالفة التعليمات أو عن الاستخدام غير السليم للمعدات.
    \item تلتزم العميلة باستخدام المعدات والأجهزة والمرافق وفق تعليمات المدربات والموظفات فقط، وعدم استخدام أي جهاز أو أداة عند وجود خلل أو دون توجيه مناسب.
    \item توافق العميلة على تعويض الاستوديو عن أي ضرر أو تكلفة أو مطالبة تنتج عن إساءة استخدام المرافق أو مخالفة الأنظمة أو التسبب في أضرار للغير أو للممتلكات.
\end{enumerate}

\section*{9. قواعد السلوك والاحترام}
\begin{enumerate}
    \item يجب احترام الموظفات والمدربات وباقي العميلات والالتزام بالسلوك اللائق في جميع الأوقات.
    \item يُمنع منعاً باتاً أي سلوك مسيء أو عدواني أو تهديدي أو متحرش أو عنصري أو غير محترم أو مخل بالآداب.
    \item يحق للاستوديو إيقاف أو تعليق أو إلغاء العضوية أو الحجز أو منع الدخول فوراً دون استرداد عند أي مخالفة للسلوك أو السياسة أو النظام الداخلي.
    \item يجب احترام العادات والتقاليد والأنظمة المحلية المعمول بها في المملكة العربية السعودية.
    \item أي إزعاج أو تعطيل لسير الحصة أو تعليمات المدربة قد يؤدي إلى استبعاد العميلة من الحصة أو من الاستوديو.
\end{enumerate}

\section*{10. الجداول والمدربات والتشغيل}
\begin{enumerate}
    \item يحق للاستوديو تعديل الجداول أو أوقات الحصص أو المدربات أو الفروع أو الخدمات أو إلغاء أو إعادة جدولة الحصص عند الحاجة ولأي أسباب تشغيلية أو فنية أو صحية أو ظروف خارجة عن الإرادة، ودون إشعار مسبق متى كان ذلك ضرورياً.
    \item في حال إلغاء الحصة من قبل الاستوديو، يحق للاستوديو إعادة الحصة إلى الرصيد أو إعادة الجدولة أو تقديم بديل مناسب وفق ما يراه مناسباً.
    \item قد تختلف ساعات العمل خلال الإجازات الرسمية أو المواسم أو شهر رمضان أو بسبب الصيانة أو الإغلاق المؤقت.
    \item إذا تم إغلاق أحد الفروع أو المرافق مؤقتاً، فقد يقوم الاستوديو بتوجيه العميلات إلى فرع آخر أو تمديد العضوية أو الرصيد أو تقديم معالجة بديلة وفق السياسة التشغيلية المعلنة.
\end{enumerate}

\section*{11. الملابس والمقتنيات الشخصية والخزائن}
\begin{enumerate}
    \item يجب ارتداء ملابس رياضية مناسبة ومحتشمة تتوافق مع طبيعة النشاط.
    \item يجب ارتداء الجوارب المانعة للانزلاق أو Grip Socks في الحصص التي يحددها الاستوديو أو في جميع الحصص إذا أعلن ذلك.
    \item يُمنع الأكل أو مضغ العلكة أو ارتداء ما قد يشكل خطراً أثناء الحصة، ويشمل ذلك المجوهرات إذا طلبت المدربة ذلك.
    \item يمنع إدخال حقائب الجيم إلى داخل قاعة التمرين إلا إذا سمح الاستوديو بذلك.
    \item توفر الخزائن للاستخدام المؤقت أثناء الزيارة فقط، ويحق للاستوديو فتحها أو إخلاؤها بعد انتهاء ساعات العمل أو عند الحاجة، دون مسؤولية عن أي محتويات متروكة.
    \item لا يتحمل الاستوديو مسؤولية فقدان أو تلف أو سرقة المقتنيات الشخصية، سواء داخل الخزائن أو خارجها.
    \item يجب استلام المفقودات خلال المدة التي يحددها الاستوديو، ويجوز التخلص من المقتنيات غير المطالب بها أو التبرع بها بعد مضي \engnum{7} أيام أو المدة التي يحددها الاستوديو، ويجوز للممتلكات ذات القيمة المنخفضة التخلص منها وفق تقدير الإدارة.
\end{enumerate}

\section*{12. التصوير والتسجيل والخصوصية}
\begin{enumerate}
    \item يمنع على العميلات والزائرات استخدام الكاميرات أو أجهزة التسجيل أو تصوير أو تسجيل الفيديو أو الصوت داخل الاستوديو أو أثناء الحصص دون موافقة كتابية مسبقة من الإدارة.
    \item يجوز للاستوديو تصوير أو تسجيل بعض الأنشطة أو الفعاليات لأغراض تشغيلية أو تسويقية أو ترويجية، ويجب على العميلة إبلاغ الاستوديو أثناء التسجيل أو قبله إذا كانت لا ترغب في الظهور في المواد الترويجية.
    \item يجمع الاستوديو ويعالج البيانات الشخصية اللازمة للتسجيل والتشغيل والتواصل وتحسين الخدمات، مع الحرص على خصوصية وسرية البيانات وفق الأنظمة المعمول بها.
    \item لا يشارك الاستوديو البيانات الشخصية مع الغير إلا عند الحاجة التشغيلية المشروعة أو بموافقة العميلة أو وفق ما يقتضيه النظام.
    \item لا يضمن الاستوديو أمان أي معلومات يتم الإفصاح عنها عبر الإنترنت بشكل مطلق، كما لا يتحمل مسؤولية سياسات الخصوصية الخاصة بالمواقع أو الخدمات التابعة للغير.
    \item إذا لم ترغب العميلة في استخدام بياناتها لإرسال الرسائل الترويجية أو العروض، فعليها إبلاغ الاستوديو بذلك.
\end{enumerate}

\section*{13. التدخين والطعام واستخدام المرافق}
\begin{enumerate}
    \item يمنع التدخين بجميع أنواعه داخل جميع مرافق الاستوديو.
    \item يمنع تناول الطعام في المناطق غير المخصصة لذلك حفاظاً على النظافة والسلامة.
    \item يجب استخدام جميع مرافق ومعدات الاستوديو بعناية ووفق التعليمات المعلنة أو تعليمات الموظفات.
\end{enumerate}

\section*{14. الأطفال والبرامج المخصصة لهم إن وجدت}
\begin{enumerate}
    \item لا يسمح باصطحاب الأطفال إلى المرافق أو الحصص غير المخصصة لهم.
    \item إذا قدم الاستوديو برامج للأطفال أو الناشئات، فللمدربات والإدارة الحق في اتخاذ ما يلزم لمعالجة السلوك المزعج أو تعليق أو إنهاء مشاركة الطفلة المسجلة عند الضرورة.
    \item يجب على ولي الأمر عدم إحضار الطفلة إذا كانت تعاني من مرض معدٍ أو حالة صحية قد تؤثر على سلامتها أو سلامة الآخرين، ويجوز للاستوديو رفض دخولها أو طلب اصطحابها إلى المنزل.
    \item يجب على ولي الأمر الإفصاح عن أي حساسية أو حالة صحية أو احتياجات خاصة قد تؤثر على مشاركة الطفلة.
    \item يتحمل أولياء الأمور أو الأوصياء مسؤولية إحضار واستلام الأطفال في الأوقات والمواقع المحددة من قبل الاستوديو.
    \item في حال وجود تصوير لبرامج الأطفال، يجب على ولي الأمر إبلاغ الاستوديو مسبقاً إذا كان لا يرغب في تصوير طفلته.
\end{enumerate}

\section*{15. الأحكام العامة}
\begin{enumerate}
    \item يحق للاستوديو تعديل هذه الشروط والأحكام أو السياسات أو الأسعار أو الخدمات أو ساعات العمل في أي وقت، ويعد استمرار استخدام الخدمات بعد التعديل موافقة على النسخة المحدثة.
    \item تخضع هذه الشروط والأحكام لأنظمة المملكة العربية السعودية، وتفسر وفقها، وتخضع أي منازعة للجهات المختصة داخل المملكة، ما لم ينص النظام على خلاف ذلك.
    \item بالموافقة على هذه الشروط أو باستخدام خدمات الاستوديو، تقر العميلة بأنها قرأت هذه الشروط وفهمتها ووافقت على الالتزام بها.
\end{enumerate}

\newpage

\selectlanguage{english}

\begin{center}
    {\Large\bfseries Slou Studio Terms \& Conditions}
\end{center}

\vspace{1em}

\section*{1. Booking and Registration}
\begin{enumerate}
    \item All bookings are subject to availability.
    \item Clients must complete the registration process and provide accurate, complete, and up-to-date information.
    \item Bookings and purchases may only be made through the studio’s official approved channels and platforms.
    \item The studio reserves the right to refuse or cancel any booking, membership, or package in the event of a breach of these Terms or for operational, security, or health-related reasons.
\end{enumerate}

\section*{2. Pricing, Payment, and VAT}
\begin{enumerate}
    \item All published prices include VAT.
    \item Prices, fees, packages, products, and services may be changed or updated at any time without prior notice.
    \item Fees must be paid in full using an approved payment method before services are used, unless the studio states otherwise.
    \item No direct payments or gifts may be given to trainers or staff in exchange for services. Clients bear full responsibility for any personal payments made outside the studio’s official channels.
\end{enumerate}

\section*{3. Memberships, Packages, and Sessions}
\begin{enumerate}
    \item All memberships, packages, and sessions are non-refundable, non-redeemable for cash, and non-exchangeable, and may not be transferred to another person unless the studio expressly agrees otherwise in writing.
    \item A single session is valid for 7 days from the date of purchase or activation, and a 10-session package is valid for 30 days from the date of purchase or activation.
    \item Sessions and packages must be used before their expiry date. Any unused sessions after expiry will be forfeited unless the studio decides otherwise.
    \item The studio reserves the right to determine the number of sessions, classes, or activities included in each membership or package according to the published plan.
    \item Only one session may be booked per person per day unless the studio states otherwise.
    \item Packages, sessions, and special offers are subject to their own published pricing and conditions.
\end{enumerate}

\section*{4. Transfers, Branch Moves, and Freezing}
\begin{enumerate}
    \item As a general rule, memberships, packages, and sessions are personal and may not be transferred to another individual.
    \item If the studio offers membership transfers, branch transfers, or internal plan changes, these will be subject to studio approval and the applicable terms and fees, including any free transfer allowance and any fee for subsequent transfers where a minimum remaining balance applies.
    \item If the studio offers freezing, the number and duration of freezes will depend on the package or plan purchased and the policy announced at the time of purchase, and may include limited freeze periods at no additional cost.
\end{enumerate}

\section*{5. Cancellations, No-Shows, Late Arrival, and Waitlists}
\begin{enumerate}
    \item Cancellations must be made at least 6 hours before the scheduled class time.
    \item Where cancellation is made within the permitted notice period, the session may be returned to the package or credit balance in accordance with the policy applicable to that booking type.
    \item Late cancellations and no-shows will result in the session being deducted or the full fee being charged, depending on the service and package, without refund.
    \item Repeated late cancellations or no-shows may result in temporary suspension of booking privileges or other booking restrictions, including a suspension period such as 3 days or longer if the studio considers it appropriate.
    \item Clients should arrive in sufficient time before class.
    \item Clients may enter up to 15 minutes after the scheduled class start time. After that, the studio may deny entry, and the booking may be treated as a no-show or late cancellation without refund.
    \item Where a waitlist feature is available, joining the waitlist may result in automatic enrollment once a spot becomes available, and the applicable cancellation policy will then apply.
    \item Clients are responsible for removing themselves from the waitlist and monitoring their bookings. If a client joins multiple waitlists or multiple sessions, they agree that sessions or charges may be applied in accordance with the policy once a booking is confirmed.
\end{enumerate}

\section*{6. Eligibility, Age, and Identification}
\begin{enumerate}
    \item Memberships and regular classes are intended for clients aged 16 and above unless a specific program states otherwise.
    \item Any program for minors or children is subject to its own announced age limits, parental or guardian consent, and any additional program-specific rules.
    \item Clients must present an ID, residence permit, or any required digital verification app or identity document if requested by the studio or required by applicable regulations.
    \item Use of a membership or booking by an unregistered or unauthorized person may result in immediate cancellation without compensation.
\end{enumerate}

\section*{7. Health, Fitness, and Medical Disclosure}
\begin{enumerate}
    \item Clients are responsible for maintaining any medical or health insurance and obtaining medical clearance where necessary, and for ensuring they are fit to participate.
    \item Clients must disclose before participation any medical condition, injury, pregnancy, recent surgery, physical restriction, or any matter that may affect safe participation.
    \item The studio may require medical approval or a medical report before allowing participation in any class or activity.
    \item Clients confirm that they are in suitable health to participate in exercise activities and have consulted a physician when appropriate, especially in the case of chronic illness, high blood pressure, diabetes, pregnancy, or other health concerns.
    \item Clients must immediately inform the trainer or studio team if they experience pain, dizziness, breathing difficulty, or any unusual symptoms before, during, or after class.
    \item The studio may stop or refuse participation temporarily or permanently if it believes this is necessary for the client’s safety or the safety of others.
    \item The studio may provide basic first aid or seek emergency medical assistance where necessary.
\end{enumerate}

\section*{8. Liability and Assumption of Risk}
\begin{enumerate}
    \item Participation in classes and activities and use of the studio’s facilities and equipment are at the client’s own risk.
    \item Clients acknowledge and accept the ordinary risks associated with physical activity and exercise, including strain, falls, injury, or health complications that may arise during or after participation.
    \item To the extent permitted by law, the studio is not liable for injuries, losses, or damages arising from participation, breach of instructions, or improper use of equipment.
    \item Clients must use all equipment, devices, and facilities only in accordance with the instructions of trainers and staff, and may not use any equipment that appears faulty or that they have not been instructed to use.
    \item Clients agree to indemnify the studio for any damage, cost, or claim resulting from misuse of facilities, breach of rules, or damage caused to other persons or property.
\end{enumerate}

\section*{9. Conduct, Respect, and Policy Enforcement}
\begin{enumerate}
    \item Clients must treat staff, trainers, and other clients with respect at all times and behave appropriately.
    \item Any abusive, aggressive, threatening, harassing, discriminatory, indecent, or otherwise inappropriate behavior is strictly prohibited.
    \item The studio may suspend or terminate a booking or membership, or deny access immediately and without refund, in the event of misconduct or violation of studio rules or policies.
    \item Clients must respect local customs, traditions, and applicable laws and regulations in the Kingdom of Saudi Arabia.
    \item Any disruption to class operations or failure to follow trainer instructions may result in removal from the class or from the studio.
\end{enumerate}

\section*{10. Schedule, Instructors, and Operations}
\begin{enumerate}
    \item The studio may amend schedules, class times, instructors, branches, or services, and may cancel or reschedule classes where necessary for operational, technical, health, or force majeure reasons, including without prior notice where required.
    \item If the studio cancels a class, it may return the session credit, reschedule the booking, or provide a suitable alternative at its discretion.
    \item Operating hours may vary during public holidays, special seasons, Ramadan, maintenance periods, or temporary closures.
    \item In the event of a temporary branch or facility closure, the studio may redirect clients to another branch or extend memberships or credits, or provide another reasonable operational solution in accordance with its announced policy.
\end{enumerate}

\section*{11. Attire, Personal Belongings, and Lockers}
\begin{enumerate}
    \item Appropriate and modest sportswear suitable for exercise must be worn.
    \item Grip socks or non-slip socks must be worn in any classes designated by the studio, or in all classes if announced by the studio.
    \item Eating, chewing gum, or wearing items that may create a safety risk during class is prohibited, including jewelry where a trainer instructs otherwise.
    \item Gym bags may not be brought into the training room unless the studio expressly permits this.
    \item Lockers are provided for temporary use during a visit only. The studio may open or clear lockers after operating hours or when necessary, without liability for contents left behind.
    \item The studio is not responsible for loss, theft, or damage to personal belongings, whether inside or outside lockers.
    \item Lost property must be collected within the period specified by the studio. Unclaimed items may be disposed of or donated after 7 days or such other period as the studio states, and low-value items may be discarded at management’s discretion.
\end{enumerate}

\section*{12. Photography, Recording, and Privacy}
\begin{enumerate}
    \item Clients and visitors may not use cameras, recording devices, or make audio or video recordings inside the studio or during classes without prior written approval from management.
    \item The studio may take photographs or recordings of activities or events for operational, marketing, or promotional purposes. Clients who do not wish to appear in such materials must notify the studio at or before registration.
    \item The studio collects and processes personal data necessary for registration, operations, communication, and service improvement, and takes privacy and confidentiality seriously in accordance with applicable laws.
    \item Personal data will not be shared with third parties except where necessary for legitimate operational purposes, with the client’s consent, or as required by law.
    \item The studio cannot guarantee the absolute security of information disclosed online and is not responsible for the privacy policies of third-party websites or services.
    \item Clients who do not wish to receive promotional messages or offers must inform the studio.
\end{enumerate}

\section*{13. Smoking, Food, and Use of Facilities}
\begin{enumerate}
    \item Smoking of any kind is strictly prohibited in all studio facilities.
    \item Food may not be consumed in areas not designated for that purpose in order to preserve hygiene and safety.
    \item All studio facilities and equipment must be used carefully and in accordance with posted or verbal instructions.
\end{enumerate}

\section*{14. Children’s Programs, If Offered}
\begin{enumerate}
    \item Children may not accompany clients into facilities or classes unless the area or class is specifically designated for them.
    \item If the studio offers children’s or youth programs, trainers and management may take appropriate action to address disruptive behavior, including suspension or termination of the enrolled child’s participation where necessary.
    \item Parents or guardians must not bring a child who has a contagious illness or a condition that may affect the child’s safety or the safety of others. The studio may refuse entry or require that the child be taken home.
    \item Parents or guardians must disclose any allergies, medical conditions, or special needs that may affect the child’s participation.
    \item Parents or legal guardians are responsible for drop-off and pick-up of children at the times and locations designated by the studio.
    \item If photography is used in children’s programs, parents or guardians must notify the studio in advance if they do not wish the child to be photographed.
\end{enumerate}

\section*{15. General Provisions}
\begin{enumerate}
    \item The studio may amend these Terms and Conditions, policies, pricing, services, or operating hours at any time. Continued use of the studio’s services after any amendment constitutes acceptance of the updated version.
    \item These Terms and Conditions are governed by the laws and regulations of the Kingdom of Saudi Arabia. Any dispute shall be subject to the competent authorities within the Kingdom unless applicable law requires otherwise.
    \item By agreeing to these Terms or using the studio’s services, the client confirms that she has read, understood, and agreed to comply with them.
\end{enumerate}

\end{document}