\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}

\title{Testing Matching Performance}
\author{Meshal Binnasban}
\date{\today}

\begin{document}
\maketitle
\newpage

\section*{Testing Matching Performance}
These tests are based on the regex enumeration code and the matching string regeneration code.

The algorithms tested are:  
1. Derivatives with simplifications.  
2. Play article's original marked approach.  
3. Shifts, a new variation developed with input-carrying marks.

The test computes the average of a fixed number of runs of each matcher for every regex–string pair.

\section{First Tests}
We start with $1{,}000{,}000$ enumerated regexes and $9$ matching strings.  
Each matcher was executed $1000$ times per regex–string pair, and the average times  
(in seconds) were recorded. The results are:

\[
\text{Derivatives} = 1.6801 \times 10^{-7}
\]

\[
\text{Play} = 2.4703 \times 10^{-7}
\]

\[
\text{Shifts} = 5.6539 \times 10^{-7}
\]

\subsection*{Summary}
The Derivatives matcher achieved the lowest average time, followed by Play, with Shifts being the slowest.  
These regexes and strings are sequentially enumerated, giving a potential advantage to the Derivatives approach.

\section{Parallel Test with Batching}
In this test, we enumerated $1{,}000{,}000$ regexes in parallel,  
using a batch size of $100{,}000$ regexes.  
For each regex, up to $9$ matching strings were generated.  
Each algorithm was run $1000$ times per regex--string pair,  
and the average times (in seconds) were recorded.

\[
\text{Derivatives} = 3.997828078658495 \times 10^{-7}
\]
\[
\text{Play} = 6.267262704770589 \times 10^{-7}
\]
\[
\text{Shifts} = 1.151973014828378 \times 10^{-6}
\]

\noindent
\textbf{Summary:} In this parallel setting, the Derivatives matcher remained the fastest,  
followed by Play, while Shifts continued to be the slowest.  
The performance gap between Derivatives and Play widened compared to earlier sequential runs,  
and Shifts was significantly slower than both.

\end{document}



\newcommand{\mycomment}[1]{}
\mycomment{

\section{Second Tests}
In this run, the regex enumeration started from $10{,}000{,}000$ and ended at $10{,}500{,}000$,  
with up to $9$ matching strings generated for each regex.  
As before, each matcher was executed $1000$ times per regex–string pair, and the average times  
(in seconds) were recorded.

\[
\text{Derivatives} = 2.2775 \times 10^{-7}
\]

\[
\text{Play} = 3.5363 \times 10^{-7}
\]

\[
\text{Shifts} = 6.5975 \times 10^{-7}
\]

\subsection*{Summary}
In this interval, the Derivatives matcher remained the fastest, followed by Play,  
while Shifts was still the slowest by a significant margin.  
The gap between Derivatives and Play widened slightly compared to the first test,  
and Shifts exhibited more than double the average time of Derivatives.
}